---
title: "ADA"
author: "Zhiqing"
date: '2022-11-16'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Introduction 
This demo will use the NHANES (National Health and Nutrition Examination Survey) dataset from 2017- 2020.03 to ask and answer these questions:  

a.Determine whether high blood pressure is associated with diabetic retinopathy or eyes affected among US Diabetic patients (45 years and above)

b.Determine whether high cholesterol level is associated with diabetic retinopathy or eyes affected among US Diabetic patients (45 years and above) 

c.Use Sex and Age as confounders/modifiers to build models.

```{r}
## Load packages and open libraries 
library(dplyr) 
library(table1)
library(tidyverse)
library(readxl)
pacman::p_load(odds.n.ends, blorr, lmtest, car,  broom, jtools) 
install.packages("foreign")
library(foreign)
```

```{r}
#select ID, Diabetic retinopathy(or eye affected) and Diabetic status from original dataset
dataset17_Q1 <- read.xport("D:/2022-Fall/Data/Project/17/P_DIQ.xpt")
dataset17_Q1 <- dataset17_Q1%>% 
  select(`SEQN`,`DIQ080`,`DIQ010`)
#select ID, Age and Sex from original dataset
dataset17_demo <- read.xport("D:/2022-Fall/Data/Project/17/P_DEMO.xpt")
dataset17_demo <- dataset17_demo%>% 
  select(`SEQN`,`RIDAGEYR`,`RIAGENDR`)
#select ID, blood pressure and blood cholestrol level from original dataset
dataset17_Q2 <- read.xport("D:/2022-Fall/Data/Project/17/P_BPQ.xpt")
dataset17_Q2 <- dataset17_Q2%>% 
  select(`SEQN`,`BPQ020`,`BPQ080`)

#merge three datasets by ID
dataset17_total <- merge(dataset17_Q1,dataset17_demo, by = "SEQN")
dataset17_total <- merge(dataset17_total,dataset17_Q2, by = "SEQN")

#Drop all NA
dataset_all <- drop_na(dataset17_total)
```

```{r}
#recode DR or eye affected
dataset_all <- dataset_all %>% 
  mutate(DIQ080_new = case_when(DIQ080 == 1 ~ 1,
                             DIQ080 == 2 ~ 0
                             ))
dataset_all$DIQ080_new<- factor(dataset_all$DIQ080_new, levels = c(0:1), labels = c("No_DR_or_Eye_affected", "DR_or_Eye_affected"))

#check this variable
table(dataset_all$DIQ080_new)
```
```{r}
#filter diabetic status, choosing those with diabetes
dataset_all <- dataset_all%>%
  filter(DIQ010 == 1 )
```

```{r}
#recode Sex
dataset_all <- dataset_all %>% 
  mutate(RIAGENDR_new = case_when(RIAGENDR == 1 ~ 1,
                             RIAGENDR == 2 ~ 0
                             ))
dataset_all$RIAGENDR_new<- factor(dataset_all$RIAGENDR_new, levels = c(0:1), labels = c("Female", "Male"))
#check this variable
table(dataset_all$RIAGENDR_new)
```

```{r}
#recode blood pressure
dataset_all <- dataset_all %>% 
  mutate(BPQ020_new = case_when(BPQ020 == 1 ~ 1,
                             BPQ020 == 2 ~ 0
                             ))
dataset_all$BPQ020_new<- factor(dataset_all$BPQ020_new, levels = c(0:1), labels = c("No", "Yes"))
#check this variable
table(dataset_all$BPQ020_new)
class(dataset_all$BPQ020_new)
```

```{r}
#recode cholesterol level
dataset_all <- dataset_all %>% 
  mutate(BPQ080_new = case_when(BPQ080 == 1 ~ 1,
                             BPQ080 == 2 ~ 0
                             ))
dataset_all$BPQ080_new<- factor(dataset_all$BPQ080_new, levels = c(0:1), labels = c("No", "Yes"))
#check this variable
table(dataset_all$BPQ080_new)
```
```{r}
#select those age 45 and over
dataset_all <- dataset_all%>%
  filter(RIDAGEYR > 45 | RIDAGEYR == 45)
```

```{r}
#recode age
class(dataset_all$RIDAGEYR)

dataset_all <- dataset_all %>% 
  mutate(RIDAGEYR_new = case_when((RIDAGEYR > 44 & RIDAGEYR< 55)  ~ 0,
                             (RIDAGEYR > 54 & RIDAGEYR< 65)  ~ 1,
                             (RIDAGEYR > 64 & RIDAGEYR<75)  ~ 2,
                             (RIDAGEYR >74 & RIDAGEYR< 85)  ~ 3,
                             ))
dataset_all$RIDAGEYR_new<- factor(dataset_all$RIDAGEYR_new, levels = c(0:3), labels = c("45-54", "55-64","65-74","75-84"))
#check this variable
table(dataset_all$RIDAGEYR_new)
```

```{r}
#drop all NA
dataset_all.clean<-na.omit(dataset_all)
```

```{r}
#rename the columns
names(dataset_all.clean)[8:12]<-c("DR_or_Eye_affected", "Sex", "High_Blood", "High_cholesterol", "Age")
```

```{r}
# make a table 1
label(dataset_all.clean$Age)<-"Age (years)"
table1(~ Age + Sex + High_Blood + High_cholesterol | DR_or_Eye_affected, overall="Total", rowlabelhead="Variable",
       data = dataset_all.clean)
```

```{r}
#perform logistic model for High_Blood and DR_or_Eye_affected
model1<- glm(DR_or_Eye_affected ~High_Blood, data=dataset_all.clean, family="binomial")
summary(model1)

odds.n.ends(model1)
```


```{r}
#Adjust sex and age, perform logistic model for High_Blood and DR_or_Eye_affected
model2<- glm(DR_or_Eye_affected ~High_Blood+Sex+Age, data=dataset_all.clean, family="binomial")
summary(model2)

odds.n.ends(model2)
```
```{r}
#perform logistic model for High_cholesterol and DR_or_Eye_affected
model3<- glm(DR_or_Eye_affected ~High_cholesterol, data=dataset_all.clean, family="binomial")
summary(model3)
odds.n.ends(model3)
```


```{r}
#Adjust sex and age, perform logistic model for High_Blood and DR_or_Eye_affected
model4<- glm(DR_or_Eye_affected ~High_cholesterol+Sex+Age, data=dataset_all.clean, family="binomial")
summary(model4)
odds.n.ends(model4)
```
```{r}
#Check the interaction term of Sex and High_Blood
model5<- glm(DR_or_Eye_affected ~High_Blood + Sex + Sex*High_Blood, data=dataset_all.clean, family="binomial")
summary(model5)

model6<- glm(DR_or_Eye_affected ~High_Blood + Sex, data=dataset_all.clean, family="binomial")
summary(model6)

lrtest(model5, model6)
```

```{r}
#Check the interaction term of Age and High_Blood
model7<- glm(DR_or_Eye_affected ~High_Blood + Age+ Age*High_Blood, data=dataset_all.clean, family="binomial")
summary(model7)

model8<- glm(DR_or_Eye_affected ~High_Blood + Age, data=dataset_all.clean, family="binomial")
summary(model8)

lrtest(model7, model8)
```


```{r}
#Check the interaction term of Sex and High_cholesterol
model9<- glm(DR_or_Eye_affected ~High_cholesterol + Sex + Sex*High_cholesterol, data=dataset_all.clean, family="binomial")
summary(model9)

model10<- glm(DR_or_Eye_affected ~High_cholesterol + Sex, data=dataset_all.clean, family="binomial")
summary(model10)

lrtest(model9, model10)
```

```{r}
# Since the likelihood ratio test is significant, so I use Sex as effect modifier
dataset_male<- dataset_all.clean%>% filter(dataset_all.clean$Sex =='Male')
dataset_female<- dataset_all.clean%>% filter(dataset_all.clean$Sex =='Female')
```


```{r}
# Perform model in Male group
model10_1<- glm(DR_or_Eye_affected ~High_cholesterol, data=dataset_male, family="binomial")
summary(model10_1)
odds.n.ends(model10_1)
```
```{r}
# Perform model in Female group
model10_2<- glm(DR_or_Eye_affected ~High_cholesterol, data=dataset_female, family="binomial")
summary(model10_2)
odds.n.ends(model10_2)
```
```{r}
# Adjust Age, and perform model in Female group
model10_2_ad<- glm(DR_or_Eye_affected ~High_cholesterol+Age, data=dataset_female, family="binomial")
summary(model10_2_ad)
odds.n.ends(model10_2_ad)
```

```{r}
#Check the interaction term of Age and High_cholesterol
model11<- glm(DR_or_Eye_affected ~High_cholesterol + Age + Age*High_cholesterol, data=dataset_all.clean, family="binomial")
summary(model11)

model12<- glm(DR_or_Eye_affected ~High_cholesterol + Age, data=dataset_all.clean, family="binomial")
summary(model12)

lrtest(model11, model12)
```